---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

---
output: github_document
---

# costcomorbidity

`costcomorbidity` is an R package for calculating the ICD-10 based Charlson
and Elixhauser, as well as the ATC-based RxRisk comorbidity indices,
derived from predicting healthcare costs.  

The package is currently under active development. Its interface and
coefficient implementation may change before the first formal release.

## Installation

`costcomorbidity` is not currently available from CRAN. The development
version can be installed from GitHub using the `pak` package:

```r
# Install pak once, if it is not already installed
install.packages("pak")

# Install costcomorbidity from GitHub
pak::pkg_install("camnys/costcomorbidity")
```

Load the package after installation:

```r
library(costcomorbidity)
```

## Citation

The cost-based comorbidity weights implemented in `costcomorbidity` were
developed and validated in:

> Nystrand, C., Johansson, N., and Blom, J. (2026). Derivation and
> Validation of Charlson, Elixhauser and RxRisk Cost-Based Comorbidity
> Indices: A Register-Based Study of 350,000 Individuals Over 11 Years.
> *PharmacoEconomics*. Accepted for publication.
> https://doi.org/10.1007/s40273-026-01652-x

If you use these indices, please cite the methodological article and report
the version of `costcomorbidity` used. Citation information and the installed
package version can be obtained with:

```r
citation("costcomorbidity")
packageVersion("costcomorbidity")
```

## Supported indices

The package supports:

- Cost-based Charlson index
- Cost-based Elixhauser index
- Cost-based RxRisk index

The resulting score variables are named:

- `costbased_charlson`
- `costbased_elixhauser`
- `costbased_rxrisk`

## RxRisk classification

RxRisk categories are identified at runtime using the `atc_pratt`
classification provided by the `coder` R package. The full Pratt mapping
is not redistributed by `costcomorbidity`.

The cost-based RxRisk coefficients distributed by `costcomorbidity`
were estimated by the package authors.

## Example

```{r example, message=FALSE}
library(costcomorbidity)

rxrisk_indicators <- identify_comorbidities(
  people = example_people,
  codes = example_codes,
  index = "rxrisk"
)

rxrisk_scores <- apply_cost_weights(
  indicators = rxrisk_indicators,
  index = "rxrisk",
  keep_indicators = FALSE
)

head(rxrisk_scores)
```
